---
title: "Longitudinal analysis of cfRNA in the maternal circulation "
subtitle: "Changes of cfRNA by gestational age"
author: |
  | [Sung Gong](https://www.obgyn.cam.ac.uk/staff/research-staff/sung-gong/){target="_blank"}
institute: |
  | Department of Obstetrics & Gynaecology
  | University of Cambridge
#date: "`r Sys.time()`"
date: "2025-04-22"
output: 
    binb::metropolis:
        slide_level: 2 
        toc: false
        fig_caption: true
        keep_tex: false
fontsize: 10pt
classoption: aspectratio=149
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
tables: true
params:
    disease: PE
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
  knitr::opts_chunk$set(cache=TRUE, echo = FALSE, message=FALSE, warning=FALSE) # with caching

  library(magrittr)
  library(lme4)
  library(robustlmm)
  library(MuMIn)
  #install.packages("MuMIn") # v1.48.11 (R>4.4.0)
  #install.packages("https://cran.r-project.org/src/contrib/Archive/MuMIn/MuMIn_1.48.4.tar.gz", type="source", repos=NULL)

  source("libs/local.R") # global settings & functions 
  source("libs/graphic.R") # Sung's graphic function

  get_r2<- function(model) {
    # Extract fixed effects and their design matrix
    fixed_effects <- fixef(model)
    design_matrix <- getME(model, "X")

    # Compute variance of fixed effects
    mu <- design_matrix %*% fixed_effects
    varF <- var(as.vector(mu))

    # Extract random effects variance
    varComps <- VarCorr(model)
    varR <- sum(sapply(varComps, function(vc) {
      if (is.matrix(vc)) {
        return(sum(diag(vc)))
      } else {
        return(vc[1]^2)
      }
    }))

    # Extract residual variance
    varRes <- sigma(model)^2

    # Marginal R-squared: variance explained by fixed effects
    R2_marginal <- varF / (varF + varR + varRes)
    # Conditional R-squared: variance explained by both fixed and random effects
    R2_conditional <- (varF + varR) / (varF + varR + varRes)

    return(c(`R2m`= R2_marginal, `R2c`= R2_conditional))
  }
```

## Longitudinal Analysis
+ Healthy samples only.
+ 12wk, 20wk, 28wk, and 36wk dataset from term dataset (i.e. Validation dataset).
+ Robust estimation of linear mixed-effect models  via `robustlmm` R package.

## Set up dataset of transcript abundancess by GA
```{r diff_case_ga_setup1}
  ###############
  # Sample Info #
  ###############
  dt.pops.sample<-fread("data/dt.samples.csv") # n=755
  dt.pops.sample[,.N,POPSID][N>4]
  dt.pops.sample[grepl("-b$",SampleID)] %>% dcast.data.table(POPSID~GA, value.var="SampleID")
  dt.pops.sample[,.(POPSID,SampleID,GA)]
  #fwrite(dt.pops.sample[,.(POPSID,SampleID,GA)], file="data/cfRNA_samples_GA.csv") # sent to Ulla

  dt.pops.sample<-dt.pops.sample[Cohort=="term" & Condition=="Control" & !SampleID %in% c("GS-B-374-UW","GS-B-374-UW-b")] # remove these two samples as both of them flagged as "failed" by illumina (it is a 28wk control sample) # n=383
  dt.pops.sample[,.N,POPSID][N>4]
  dt.pops.sample[,.N,POPSID][N<4]
  dt.pops.sample[POPSID==1953] %>% dcast.data.table(POPSID~GA, value.var="SampleID")
  dt.pops.sample[,.(POPSID,SampleID,GA)] # n=383
  dt.pops.sample[,.N,.(POPSID)] # n=96

  dt.foo<-data.table(read_excel("data/cfRNA_samples_GA_exact.xlsx"))[,.(SampleID,GAwk=GA_exact)] # from Ulla
  dt.pops.sample<-merge(dt.pops.sample, dt.foo) # n=383
  dt.pops.sample$POPSID<-factor(dt.pops.sample$POPSID)

  #####################
  ## Based on logCPM ##
  #####################
  load("RData/dt.cpmZ.term.POPS-2022.GRCh38.88.RData") # dt.cpmZ.term (term)

  xtabs(~GA+Condition,dt.cpmZ.term)
  dt.cpmZ.term[,.(.N,N_sample=length(unique(SampleID)),N_gene=length(unique(geneName))),.(GA,Condition)]

```

## Build the models using quaratic term (i.e. GA^2) and linear term (GA)
+ Linear and non-linear models via `lme4::lmer`
+ Linear and non-linear models via `robustlmm::rlmer`
```{r rlmer_quaratic1}
  dt.logCPM<-merge(
                dt.cpmZ.term[Condition=="Control"],
                dt.pops.sample[,.(SampleID,POPSID,GAwk)],
                by="SampleID"
                )

  dt.logCPM$GAwk %>% summary
  dt.logCPM[,rGAwk:=(GAwk-4)/8] # to make 12wk(1), 20wk(2), 28wk(3), and 36wk(4)
  dt.logCPM$rGAwk %>% summary
  dl.logCPM<-split(dt.logCPM, dt.logCPM$geneName)
  dl.logCPM %>% length # n=15150
  rm(dt.logCPM)

  if(file.exists("RData/li.lmer.RData")){
    load("RData/li.lmer.RData")
  }else{
    #
    li.lmer<-
    lapply(dl.logCPM, function(dt.foo){
      my.gene<-unique(dt.foo[["geneName"]])
      message(my.gene)

      if(F){
        ## 1a. Mixed Effect Model via lmer4 (non-linear)
        system.time(my.lmer <- lme4::lmer(logCPM ~ 1 + rGAwk + I(rGAwk^2) + (1 | POPSID), data = dt.foo))
        #my.lmer %>% summary
        my.r2<-MuMIn::r.squaredGLMM(my.lmer)

        dt.lmer.predict <- 
          my.lmer %>% ggeffects::ggpredict(terms = list(rGAwk= seq(min(dt.foo$rGAwk), max(dt.foo$rGAwk), length = 20)), verbose = FALSE) %>% data.table %>% setnames("x","rGAwk") %>% setnames("predicted","logCPM")
        dt.lmer.predict[,GAwk:=rGAwk*8 +4][,group:=NULL][,rGAwk:=NULL] %>% setcolorder("GAwk")

        ## 2a. Robust Mixed Effect Model via robustlmm (non-linear)
        system.time(my.rlmer <- robustlmm::rlmer(logCPM ~ 1 + rGAwk + I(rGAwk^2) + (1 | POPSID), data = dt.foo))
        my.rlmer %>% summary
        my.r2<-get_r2(my.rlmer) %>% t 

        if(F){
          var.fix <- insight::get_variance_fixed(my.rlmer)
          var.ran <- insight::get_variance_random(my.rlmer)
          var.res <- insight::get_variance_residual(my.rlmer)
          (R2m = var.fix/(var.fix+var.ran+var.res))
          #var.fixed
          (R2c = (var.fix+var.ran)/(var.fix+var.ran+var.res))

          robustlmm::getME(my.rlmer, "w_e") # robust weight with the observation
          robustlmm::getME(my.rlmer, "w_b") # robust weight with the random effect 
        }

        ## 2b. Robust Mixed Effect Model via robustlmm (linear)
        my.rlmer.linear <- rlmer(logCPM ~ 1 + rGAwk + (1 | POPSID), data = dt.foo)
        my.rlmer.linear  %>% summary
      }

      #####################################################
      ## 1a. Mixed Effect Model via lmerTest (non-linear) #
      #####################################################
      my.lmer.GA2 <- lmerTest::lmer(logCPM ~ 1 + rGAwk + I(rGAwk^2) + (1 | POPSID), data = dt.foo)
      dt.predict.GA2<- 
        my.lmer.GA2 %>% ggeffects::ggpredict(terms = list(rGAwk= seq(min(dt.foo$rGAwk), max(dt.foo$rGAwk), length = 20)), verbose = FALSE) %>% data.table %>% setnames("x","rGAwk") %>% setnames("predicted","logCPM")
      dt.predict.GA2[,GAwk:=rGAwk*8 +4][,group:=NULL][,rGAwk:=NULL] %>% setcolorder("GAwk")

      mat.coef<-lmerTest:::get_coefmat(my.lmer.GA2)
      my.coeff.GA2<-mat.coef[2:3,1,drop=F] %>% t
      my.pval.GA2<-mat.coef[2:3,5,drop=F] %>% t
      my.r2.GA2<-MuMIn::r.squaredGLMM(my.lmer.GA2)

      dt.lmer.GA2<-data.table(my.gene,my.r2.GA2,my.coeff.GA2, my.pval.GA2) %>% setnames(c("gene","r2m","r2c","coef.GA","coef.GA2","p.GA","p.GA2"))
      #list(`stat`=dt.lmer.GA2,`predict`=dt.predict.GA2)

      ##############################################
      ## 1b. Mixed Effect Model via lmer4 (linear) #
      ##############################################
      my.lmer.GA <- lme4::lmer(logCPM ~ 1 + rGAwk + (1 | POPSID), data = dt.foo)
      dt.predict.GA <- 
        my.lmer.GA %>% ggeffects::ggpredict(terms = list(rGAwk= seq(min(dt.foo$rGAwk), max(dt.foo$rGAwk), length = 20)), verbose = FALSE) %>% data.table %>% setnames("x","rGAwk") %>% setnames("predicted","logCPM")
      dt.predict.GA[,GAwk:=rGAwk*8 +4][,group:=NULL][,rGAwk:=NULL] %>% setcolorder("GAwk")

      mat.coef<-lmerTest:::get_coefmat(my.lmer.GA)
      my.coeff.GA<-mat.coef[2:2,1,drop=F] %>% t
      my.pval.GA<-mat.coef[2:2,5,drop=F] %>% t
      my.r2.GA<-MuMIn::r.squaredGLMM(my.lmer.GA)

      dt.lmer.GA<-data.table(my.gene,my.r2.GA,my.coeff.GA,NA,my.pval.GA,NA) %>% setnames(c("gene","r2m","r2c","coef.GA","coef.GA2","p.GA","p.GA2"))
      #list(`stat`=dt.lmer.GA,`predict`=dt.predict.GA)

      ###############
      ## Anova test #
      ###############
      my.anova<-anova(my.lmer.GA2, my.lmer.GA)

      ## Summary
      dt.lmer<-
        rbind(
        cbind("Model"="GA2",dt.lmer.GA2),
        cbind("Model"="GA",dt.lmer.GA)
        )[,p.anova:=my.anova[["Pr(>Chisq)"]][2]] %>% setcolorder("gene")

      list(`stat`=dt.lmer, `predict`=dt.predict.GA2, `predict.GA`=dt.predict.GA)
    })
    save(li.lmer, file="RData/li.lmer.RData")
  }
```

## Strength of association via R^2
```{r r2_all1}
  li.lmer %>% length

  dt.R2<-
    (lapply(li.lmer, function(li.model) li.model[["stat"]]) %>% rbindlist)[order(-r2m)]
  dt.R2[,.N,gene][N<2]

  dt.R2[gene=="CSHL1"]

  ## GA2
  dt.R2.GA2<-dt.R2[Model=="GA2"]
  dt.R2.GA2[,p.combined:=pchisq(-2 * (log(p.GA) + log(p.GA2)), 2 * 2, lower.tail = FALSE)]
  dt.R2.GA2[,BH.GA:=p.adjust(p.GA,"BH")]
  dt.R2.GA2[,BH.GA2:=p.adjust(p.GA2,"BH")]
  dt.R2.GA2[,BH.combined:=p.adjust(p.combined,"BH")]
  dt.R2.GA2[,BH.anova:=p.adjust(p.anova,"BH")]
  fwrite(dt.R2.GA2, file="Tables/Suppl/SI.Table.10.lmer.GA2.csv")

  # GA
  dt.R2.GA<-dt.R2[Model=="GA"]
  dt.R2.GA[,BH.GA:=p.adjust(p.GA,"BH")]
  dt.R2.GA[,BH.anova:=p.adjust(p.anova,"BH")]
  fwrite(dt.R2.GA, file="Tables/Suppl/SI.Table.11.lmer.GA.csv")

  if(F){
    ##############
    ## GA2 model #
    ##############
    dt.R2.GA2[order(coef.GA)]
    dt.R2.GA2[order(coef.GA2)]

    dt.R2.GA2[order(-r2m)]
    dt.R2.GA2[order(-r2c)]

    dt.R2.GA2[order(-r2m)][1:100][coef.GA<0]

    dt.R2.GA2[order(p.GA)]
    dt.R2.GA2[order(p.GA2)]
    dt.R2.GA2[order(p.combined)]

    dt.R2.GA2[BH.GA<0.05][order(p.GA)]
    dt.R2.GA2[BH.GA2<0.05][order(p.GA2)]
    dt.R2.GA2[BH.combined<0.05][order(p.combined)]#[coef.GA<0]

    dt.R2.GA2[BH.combined<0.05][order(p.combined)][coef.GA<0]


    dt.R2.GA2[r2m>=0.5][order(-r2m)]
    dt.R2.GA2[r2c>=0.5][order(-r2c)]#[1:50,1:5]

    dt.R2.GA2[gene %in% c("LEP","PAPPA2","TTPAL")]

    #############
    ## GA model #
    #############
    dt.R2.GA[order(-r2m)][1:10]

    dt.R2.GA[order(-r2m)][1:100][coef.GA<0]
    dt.R2.GA[order(-r2m)][1:50][coef.GA>0]
    dt.R2.GA[BH.GA<0.05][order(p.GA)][coef.GA<0][1:10]
    dt.R2.GA[BH.GA<0.05][order(p.GA)][coef.GA<0]

    ###########
    ## Anova ##
    ###########
    dt.R2.GA2[BH.anova<0.05][order(p.anova)][1:10]
    dt.R2.GA[BH.anova<0.05][order(p.anova)][1:10]
  }

```

## Plot logCPM by GA
```{r rlmer_csh1_plot}
targets<-c(
  dt.R2.GA2[order(-r2m)][["gene"]][1:100],
  dt.R2.GA2[order(p.combined)][["gene"]][1:100],
  dt.R2.GA[BH.GA<0.05][order(p.GA)][coef.GA<0][["gene"]][1:100]
  ) %>% unique
#
  li.plots<-lapply(dl.logCPM[targets], function(dt.foo){
      my.gene<-unique(dt.foo[["geneName"]])
      message(my.gene)

      r2m<-li.lmer[[my.gene]][["stat"]][Model=="GA2"][["r2m"]] %>% round(3)
      r2c<-li.lmer[[my.gene]][["stat"]][Model=="GA2"][["r2c"]] %>% round(3)
      #my.title=paste0(my.gene," (r2m:",r2m,", r2c:",r2c,")")
      my.title=paste0(my.gene," (r2m:",r2m,")")

      my.title.y<-latex2exp::TeX(r"($\textbf{log_2 CPM}$)")
  
      ggplot(dt.foo, aes(x = GAwk, y = logCPM)) +
      geom_line(aes(group = POPSID), linewidth = .5, col="grey80",alpha=.7) +
      geom_point(aes(color = as.factor(GA)), size=2, alpha=.7) +
      labs(subtitle=my.title, x = "Gestational Age (weeks)", y=my.title.y, color = "GA (week)") +
      geom_ribbon(data=li.lmer[[my.gene]][["predict"]], aes(ymin=conf.low, ymax=conf.high), alpha=0.6, fill = "grey40") +
      geom_line(data=li.lmer[[my.gene]][["predict"]], linewidth=1.5, col="blue",alpha=0.8) + 
      scale_x_continuous(breaks=c(12,20,28,36)) +
      scale_color_manual(values=cbPalette2) +
      theme_Publication() +
      theme(legend.position="none")
  })

  # Top 100 by R2M
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.pdf'))
  cairo_pdf(filename=file.name1, width=14, height=9, onefile=T)
  grid.text(label=paste0("Supplementary Figure 4.\nTop 100 cfRNAs\nby the coefficient of determination"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-r2m)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)

  # Top 100 by p.combined
  grid.newpage()
  grid.text(label=paste0("Supplementary Figure 5.\nTop 100 cfRNAs\nby the combined p-values"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(p.combined)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)

  # Top 100 by negative coefficient in linear mixed effect model (GA only)
  grid.newpage()
  grid.text(label=paste0("Supplementary Figure 6.\nTop 100 cfRNAs\nsignificantly decreasing\n(negative coefficient)\nby gestational age"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA[BH.GA<0.05][order(p.GA)][coef.GA<0][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  if(F){
  ##
  # Top 100 by R2M
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.r2m.pdf'))
  #pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  cairo_pdf(filename=file.name1, width=14, height=9, onefile=T)
  grid.text(label=paste0("Supplementary Figure 3.\nTop 100 cfRNAs\nby the coefficient of determination"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-r2m)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by p.combined
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.p.combined.pdf'))
  cairo_pdf(filename=file.name1, width=14, height=9, onefile=T)
  grid.text(label=paste0("Supplementary Figure 4.\nTop 100 cfRNAs\nby the combined p-values"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(p.combined)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by negative coefficient in linear mixed effect model (GA only)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.decreasing.pdf'))
  cairo_pdf(filename=file.name1, width=14, height=9, onefile=T)
  grid.text(label=paste0("Supplementary Figure 5.\nTop 100 cfRNAs\nsignificantly decreasing\n(negative coefficient)\nby gestational age"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.6,just="center")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA[BH.GA<0.05][order(p.GA)][coef.GA<0][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()
  }

  if(F){
  # Bottom 100 by negative coefficient in linear mixed effect model (GA only)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.bottom100.decreasing.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA[BH.GA<0.05][order(-p.GA)][coef.GA<0][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by R2C
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.r2c.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  grid.text(label=paste0("Longitudinal analysis of cfRNA\nfrom maternal blood\n(Top 100 genes by R2C)"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.7,just="center")
  grid.text(label=paste("Sungsam Gong\n<ssg29 @ cam.ac.uk>\n",Sys.Date()),gp=gpar(fontsize=30,fontface="bold"),x=0.8,y=.2,just="right")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-r2c)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by abs(coef.GA)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.coefGA.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  grid.text(label=paste0("Longitudinal analysis of cfRNA\nfrom maternal blood\n(Top 100 genes by coef(GA))"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.7,just="center")
  grid.text(label=paste("Sungsam Gong\n<ssg29 @ cam.ac.uk>\n",Sys.Date()),gp=gpar(fontsize=30,fontface="bold"),x=0.8,y=.2,just="right")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-abs(coef.GA))][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by pos(coef.GA)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.pos.coefGA.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  grid.text(label=paste0("Longitudinal analysis of cfRNA\nfrom maternal blood\n(Top 100 genes by coef(GA))>0"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.7,just="center")
  grid.text(label=paste("Sungsam Gong\n<ssg29 @ cam.ac.uk>\n",Sys.Date()),gp=gpar(fontsize=30,fontface="bold"),x=0.8,y=.2,just="right")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-coef.GA)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by neg(coef.GA)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.neg.coefGA.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  grid.text(label=paste0("Longitudinal analysis of cfRNA\nfrom maternal blood\n(Top 100 genes by coef(GA))<0"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.7,just="center")
  grid.text(label=paste("Sungsam Gong\n<ssg29 @ cam.ac.uk>\n",Sys.Date()),gp=gpar(fontsize=30,fontface="bold"),x=0.8,y=.2,just="right")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(coef.GA)][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()

  # Top 100 by abs(coef.GA2)
  file.name1<-file.path('Figures/Suppl',paste0('cfRNA.Suppl.Data.Fig.logCPM_by_GA_lmer.top100.coefGA2.pdf'))
  pdf(file=file.name1, width=14, height=9,title="Longitudinal analysis via mixed effect model")
  grid.text(label=paste0("Longitudinal analysis of cfRNA\nfrom maternal blood\n(Top 100 genes by coef(GA2))"),gp=gpar(fontsize=50,fontface="bold"),x=.5,y=.7,just="center")
  grid.text(label=paste("Sungsam Gong\n<ssg29 @ cam.ac.uk>\n",Sys.Date()),gp=gpar(fontsize=30,fontface="bold"),x=0.8,y=.2,just="right")
  gridExtra::marrangeGrob(li.plots[dt.R2.GA2[order(-abs(coef.GA2))][["gene"]][1:100]],nrow=2,ncol=4,top=NULL)
  dev.off()
  }

```

## Plot an example
```{r rlmer_csh1_plot_foo1, eval=F}
  my.gene<-"CSHL1"
  my.gene<-"TTPAL"

  dt.foo<-dl.logCPM[[my.gene]]
  system.time(my.lmer <- lme4::lmer(logCPM ~ 1 + rGAwk + I(rGAwk^2) + (1 | POPSID), data = dt.foo))
  system.time(my.lmer.test <- lmerTest::lmer(logCPM ~ 1 + rGAwk + I(rGAwk^2) + (1 | POPSID), data = dt.foo))

  lmerTest:::get_coefmat(my.lmer.test)
  MuMIn::coeffs(my.lmer.test)
  MuMIn::r.squaredGLMM(my.lmer.test)

  # Predictions for the robust model
  dt.lmer.predict <- ggeffects::ggpredict(li.models[[my.gene]][["lmer"]],
                      terms = list(rGAwk= seq(min(dl.logCPM[[my.gene]]$rGAwk), max(dl.logCPM[[my.gene]]$rGAwk), length = 30)),
                      verbose = FALSE) %>% data.table %>% setnames("x","rGAwk") %>% setnames("predicted","logCPM")
  dt.lmer.predict[,GAwk:=rGAwk*8 +4][,group:=NULL][,rGAwk:=NULL] %>% setcolorder("GAwk")



  p1<-plot(li.models[[my.gene]][["lmer"]], which=1)

  p.all<-
  ggplot(dl.logCPM[[my.gene]], aes(x = GAwk, y = logCPM)) +
  geom_line(aes(group = POPSID), linewidth = .5, col="grey80",alpha=.7) +
  geom_point(aes(color = as.factor(GA)), size=2, alpha=.7) +
  labs(x = "Gestational Age", color = "GA (week)") +
  theme(legend.position = "top") +
  geom_ribbon(data=dt.lmer.predict, aes(ymin=conf.low, ymax=conf.high), alpha=0.6, fill = "grey40") +
  geom_line(data=dt.lmer.predict, size=1.5, col="blue",alpha=0.8) + 
  scale_x_continuous(breaks=c(12,20,28,36)) +
  scale_color_manual(values=cbPalette2) +
  theme_Publication() +
  theme(legend.position="none")

  #pdf(file="Figures/RLMM.CSH1.by.GA.pdf")
  cairo_pdf(filename=file.path("Figures",paste0("RLMM.",my.gene,".by.GA.pdf")), onefile=T)
  cowplot::plot_grid(p1)
  p2 #+ theme_Publication()
  #p3 #+ theme_Publication()
  p.all
  dev.off()
```
